---
title: "Biostat 203B Homework 2"
subtitle: Due Feb 6 @ 11:59PM
author: Michael Markot
output: 
  html_document:
    toc: true
    toc_depth: 4 
---

Display machine information for reproducibility:
```{r}
sessionInfo()
```

```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE)
library(tidyverse)
library(data.table)
library(lubridate)
```

```{r}
os <- sessionInfo()$running
if (str_detect(os, "Linux")) {
  mimic_path <- "/mnt/mimiciv/1.0"
} else if (str_detect(os, "macOS")) {
  mimic_path <- "/Users/huazhou/Documents/Box Sync/MIMIC/mimic-iv-1.0"
}

mimic_path <- "/mnt/mimiciv/1.0"
```

In this exercise, we use tidyverse (ggpot2, dplyr, etc) to explore the [MIMIC-IV](https://mimic.mit.edu/docs/iv/) data introduced in [homework 1](https://ucla-biostat-203b.github.io/2022winter/hw/hw1/hw1.html) and to build a cohort of ICU stays.

```{r}
# tree -s -L 2 /Users/huazhou/Documents/Box\ Sync/MIMIC/mimic-iv-1.0
system(str_c("tree -s -L 2 ", shQuote(mimic_path)), intern = TRUE)
```

## Q1. `read.csv` (base R) vs `read_csv` (tidyverse) vs `fread` (data.table)

There are quite a few utilities in R for reading plain text data files. Let us test the speed of reading a moderate sized compressed csv file, `admissions.csv.gz`, by three programs: `read.csv` in base R, `read_csv` in tidyverse, and `fread` in the popular data.table package. 

```{r, eval =T}
system.time(read.csv("/mnt/mimiciv/1.0/core/admissions.csv.gz"))
system.time(read_csv("/mnt/mimiciv/1.0/core/admissions.csv.gz"))
system.time(fread("/mnt/mimiciv/1.0/core/admissions.csv.gz"))
```

Which function is fastest? Is there difference in the (default) parsed data types? (Hint: R function `system.time` measures run times.)

**Solution:** The 'fread' function is fastest, followed clsoely by 'read_csv'. 
'read.csv' was the slowest by more than a factor of ten. 'read_csv' classifies 
the time variables as date time data types whereas 'read.csv' classifies these 
by default as numeric.

For later questions, we stick to the tidyverse.

## Q2. ICU stays

`icustays.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/icustays/>) contains data about Intensive Care Units (ICU) stays. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/icustays.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `icustatys.csv.gz` as a tibble `icustays_tble`. 

```{r}
icustays_tble <- read_csv("/mnt/mimiciv/1.0/icu/icustays.csv.gz")
```

2. How many unique `subject_id`? Can a `subject_id` have multiple ICU stays?

```{r}
length(unique(icustays_tble$subject_id))
```
**Solution:** There are 53,150 unique subject ID's. There are more total 
observations than unique subject ID's so a subject ID must be able to have 
multiple stays.

3. For each `subject_id`, let's only keep the first ICU stay in the tibble `icustays_tble`.

```{r}
icustays_tble <- icustays_tble %>% 
                   group_by(subject_id) %>% 
                   arrange(intime) %>% 
                   distinct(subject_id, .keep_all = TRUE)
```

## Q3. `admission` data

Information of the patients admitted into hospital is available in `admissions.csv.gz`. See <https://mimic.mit.edu/docs/iv/modules/core/admissions/> for details of each field in this file. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/core/admissions.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `admissions.csv.gz` as a tibble `admissions_tble`.

```{r}
admissions_tble <- read_csv("/mnt/mimiciv/1.0/core/admissions.csv.gz")
```

2. Let's only keep the admissions that have a match in `icustays_tble` according to `subject_id` and `hadmi_id`.

```{r}
admissions_tble <- admissions_tble %>% 
  semi_join(icustays_tble, by = "subject_id", "hadm_id")
```

3. Summarize the following variables by graphics. 

- admission year  

```{r}
admissions_tble$admit_year <- year(admissions_tble$admittime)
ggplot(data = admissions_tble) +
  geom_bar(mapping = aes(admit_year))
```

- admission month  

```{r}
admissions_tble$admit_month <- month(admissions_tble$admittime, label = TRUE)
ggplot(data = admissions_tble) +
  geom_bar(mapping = aes(admit_month))
```

- admission month day  

```{r}
admissions_tble$admit_mday <- mday(admissions_tble$admittime)
ggplot(data = admissions_tble) +
  geom_bar(mapping = aes(admit_mday))
```

- admission week day  

```{r}
admissions_tble$admit_wday <- wday(admissions_tble$admittime, label = TRUE)
ggplot(data = admissions_tble) +
  geom_bar(mapping = aes(admit_wday))
```

- admission hour (anything unusual?)  

```{r}
admissions_tble$admit_hour <- hour(admissions_tble$admittime)
ggplot(data = admissions_tble) +
  geom_bar(mapping = aes(admit_hour))
```


**Solution:** There is a large spike in admissions at 7am compared to the hours 
preceding and following it. This could be when most people wake up and decide 
they need to go to the hospital, when there is a staffing increase at the 
hospital for the beginning of the work day, or another unknown factor.

## Q4. `patients` data

Patient information is available in `patients.csv.gz`. See <https://mimic.mit.edu/docs/iv/modules/core/patients/> for details of each field in this file. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/core/patients.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `patients.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/core/patients/>) as a tibble `patients_tble` and only keep the patients who have a match in `icustays_tble` (according to `subject_id`).

```{r}
patients_tble <- read_csv("/mnt/mimiciv/1.0/core/patients.csv.gz")
```

```{r}
patients_tble <- patients_tble %>% 
  semi_join(icustays_tble, by = "subject_id")
```

2. Summarize variables `gender` and `anchor_age`, and explain any patterns you see.

```{r}
patients_tble$gender <- as.factor(patients_tble$gender)
summary.factor(patients_tble$gender)
summary(patients_tble$anchor_age)
```

**Solution:** There are 56% males comapred to females and the age seems to be 
consist of many more older people than younger people with the mean and median 
being over 60. This makes sense since older people are more likly to need the 
hospital than younger peolpe in general. Also the minimum age is 18 so we can 
assume that children were excluded from this data.

## Q5. Lab results

`labevents.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/hosp/labevents/>) contains all laboratory measurements for patients. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/hosp/labevents.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

`d_labitems.csv.gz` is the dictionary of lab measurements. 

```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/hosp/d_labitems.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Find how many rows are in `labevents.csv.gz`.

```{bash eval = T}
wc -l "/mnt/mimiciv/1.0/hosp/labevents.csv.gz"
```

**Solution:** There are 8,398,755 rows in 'labevents.csv.gz'.


2. We are interested in the lab measurements of creatinine (50912), potassium (50971), sodium (50983), chloride (50902), bicarbonate (50882), hematocrit (51221), white blood cell count (51301), glucose (50931), magnesium (50960), and calcium (50893). Retrieve a subset of `labevents.csv.gz` only containing these items for the patients in `icustays_tble` as a tibble `labevents_tble`. 

    Hint: `labevents.csv.gz` is a data file too big to be read in by the `read_csv` function in its default setting. Utilize the `col_select` and `lazy` options in the `read_csv` function to reduce the memory burden.

```{r}
labpath <- "/mnt/mimiciv/1.0/hosp/labevents_filtered_itemid.csv.gz"
labevents_tble <- read_csv(labpath)
```

```{r}
labevents_tble <- labevents_tble %>% 
  semi_join(icustays_tble, by = "subject_id")
```

3. Further restrict `labevents_tble` to the first lab measurement during the ICU stay. 

```{r warning = FALSE}
labevents_tble <- labevents_tble %>% 
  group_by(subject_id, charttime) %>% 
  pivot_wider(id_cols = c(subject_id, charttime), 
    names_from = itemid, values_from = valuenum)
                                 
                     
```


```{r}
labevents_tble <- labevents_tble %>% 
                     group_by(subject_id) %>% 
                     arrange(charttime) %>% 
                     distinct(subject_id, .keep_all = TRUE)
```

```{r}
labevents_tble <- labevents_tble %>% 
  mutate(`50882` = ifelse(`50882` == "NULL", NA, unlist(`50882`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50902` = ifelse(`50902` == "NULL", NA, unlist(`50902`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50912` = ifelse(`50912` == "NULL", NA, unlist(`50912`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50931` = ifelse(`50931` == "NULL", NA, unlist(`50931`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50971` = ifelse(`50971` == "NULL", NA, unlist(`50971`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50983` = ifelse(`50983` == "NULL", NA, unlist(`50983`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`51221` = ifelse(`51221` == "NULL", NA, unlist(`51221`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`51301` = ifelse(`51301` == "NULL", NA, unlist(`51301`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50893` = ifelse(`50893` == "NULL", NA, unlist(`50893`)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(`50960` = ifelse(`50960` == "NULL", NA, unlist(`50960`)) %>% 
           as.numeric)
```


4. Summarize the lab measurements by appropriate numerics and graphics. 

```{r}
names(labevents_tble) <- c('subject_id', 'charttimelabs', 'bicarbonate', 
                       'chloride', 'creatinine', 'glucose', 'potassium', 
                       'sodium', 'hematocrit', 'white_blood_cell_count', 
                       'calcium','magnesium')
```

```{r}
labevents_tble <- labevents_tble %>% 
  mutate(creatinine = ifelse(creatinine > 100, NA, unlist(creatinine)) %>% 
           as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(white_blood_cell_count = ifelse(white_blood_cell_count > 200, NA, 
    unlist(white_blood_cell_count)) %>% 
    as.numeric)

labevents_tble <- labevents_tble %>% 
  mutate(white_blood_cell_count = ifelse(white_blood_cell_count > 200, NA, 
    unlist(white_blood_cell_count)) %>% 
    as.numeric)

```

```{r}
summary(labevents_tble[3:12])
```

```{r warning = FALSE}
mbicarbonate <- mean(labevents_tble$bicarbonate, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(bicarbonate), fill = "light blue") +
  geom_vline(xintercept = mbicarbonate, size = 1, color = "red")

mchloride <- mean(labevents_tble$chloride, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(chloride), fill = "light blue") +
  geom_vline(xintercept = mchloride, size = 1, color = "red")

mcreatinine <- mean(labevents_tble$creatinine, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(creatinine), fill = "light blue") +
  geom_vline(xintercept = mcreatinine, size = 1, color = "red")

mglucose <- mean(labevents_tble$glucose, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(glucose), fill = "light blue") +
  geom_vline(xintercept = mglucose, size = 1, color = "red")

mpotassium <- mean(labevents_tble$potassium, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(potassium), fill = "light blue") +
  geom_vline(xintercept = mpotassium, size = 1, color = "red")

msodium <- mean(labevents_tble$sodium, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(sodium), fill = "light blue") +
  geom_vline(xintercept = msodium, size = 1, color = "red")

mhematocrit <- mean(labevents_tble$hematocrit, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(hematocrit), fill = "light blue") +
  geom_vline(xintercept = mhematocrit, size = 1, color = "red")

mwhite_blood_cell_count <- mean(labevents_tble$white_blood_cell_count, 
                                na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(white_blood_cell_count), fill = "light blue") +
  geom_vline(xintercept = mwhite_blood_cell_count, size = 1, color = "red")

mcalcium <- mean(labevents_tble$calcium, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(calcium), fill = "light blue") +
  geom_vline(xintercept = mcalcium, size = 1, color = "red")

mmagnesium <- mean(labevents_tble$magnesium, na.rm = TRUE)
ggplot(data = labevents_tble) +
  geom_density(mapping = aes(magnesium), fill = "light blue") +
  geom_vline(xintercept = mmagnesium, size = 1, color = "red")
```

## Q6. Vitals from charted events

`chartevents.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/chartevents/>) contains all the charted data available for a patient. During their ICU stay, the primary repository of a patient’s information is their electronic chart. The `itemid` variable indicates a single measurement type in the database. The `value` variable is the value measured for `itemid`. The first 10 lines of `chartevents.csv.gz` are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/chartevents.csv.gz")), 
    " | head"), 
  intern = TRUE
)
```
`d_items.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/d_items/>) is the dictionary for the `itemid` in `chartevents.csv.gz`. 
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/d_items.csv.gz")), 
    " | head"), 
  intern = TRUE
)
```

1. We are interested in the vitals for ICU patients: heart rate (220045), mean non-invasive blood pressure (220181), systolic non-invasive blood pressure (220179), body temperature in Fahrenheit (223761), and respiratory rate (220210). Retrieve a subset of `chartevents.csv.gz` only containing these items for the patients in `icustays_tble` as a tibble `chartevents_tble`.

```{r}
chartpath <- "/mnt/mimiciv/1.0/icu/chartevents_filtered_itemid.csv.gz"
chartevents_tble <- read_csv(chartpath)
```

```{r}
chartevents_tble <- chartevents_tble %>% 
  semi_join(icustays_tble, by = "subject_id")
```

```{r warning = FALSE}
chartevents_tble <- chartevents_tble %>% 
  group_by(subject_id, charttime) %>% 
  pivot_wider(id_cols = c(subject_id, charttime), 
    names_from = itemid, values_from = valuenum)
```



2. Further restrict `chartevents_tble` to the first vital measurement during the ICU stay. 

```{r}
chartevents_tble <- chartevents_tble %>% 
                     group_by(subject_id) %>% 
                     arrange(charttime) %>% 
                     distinct(subject_id, .keep_all = TRUE)
```

```{r}
chartevents_tble <- chartevents_tble %>% 
  mutate(`220179` = ifelse(`220179` == "NULL", NA, unlist(`220179`)) %>% 
           as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(`220181` = ifelse(`220181` == "NULL", NA, unlist(`220181`)) %>% 
           as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(`220045` = ifelse(`220045` == "NULL", NA, unlist(`220045`)) %>% 
           as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(`220210` = ifelse(`220210` == "NULL", NA, unlist(`220210`)) %>% 
           as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(`223761` = ifelse(`223761` == "NULL", NA, unlist(`223761`)) %>% 
               as.numeric)
```

3. Summarize these vital measurements by appropriate numerics and graphics. 

```{r}
names(chartevents_tble) <- c('subject_id', 'charttime', 
                             'systolic_non_invasive_blood_pressure', 
                       'mean_non_invasive_blood_pressure', 'heart_rate', 
                       'respiratory_rate', 'body_temperature_in_Fahrenheit')
```

```{r}

chartevents_tble <- chartevents_tble %>% 
  mutate(systolic_non_invasive_blood_pressure = 
           ifelse(systolic_non_invasive_blood_pressure > 300, NA, 
                  unlist(systolic_non_invasive_blood_pressure)) %>% 
         as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(mean_non_invasive_blood_pressure = 
           ifelse(mean_non_invasive_blood_pressure > 300, NA, 
                  unlist(mean_non_invasive_blood_pressure)) %>% 
         as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(heart_rate = ifelse(heart_rate > 300, NA, unlist(heart_rate)) %>% 
           as.numeric)

chartevents_tble <- chartevents_tble %>% 
  mutate(body_temperature_in_Fahrenheit = 
           ifelse(body_temperature_in_Fahrenheit < 70, NA, 
                  unlist(body_temperature_in_Fahrenheit)) %>% 
         as.numeric)
```

```{r}
summary(chartevents_tble[3:7])
```

```{r warning = FALSE}
msniBP <- mean(chartevents_tble$systolic_non_invasive_blood_pressure,
               na.rm = TRUE)
ggplot(data = chartevents_tble) +
  geom_density(mapping = aes(systolic_non_invasive_blood_pressure), 
               fill = "light green") +
  geom_vline(xintercept = msniBP, size = 1, color = "red") +
  xlim(0, 300)

mmniBP <- mean(chartevents_tble$mean_non_invasive_blood_pressure, na.rm = TRUE)
ggplot(data = chartevents_tble) +
  geom_density(mapping = aes(mean_non_invasive_blood_pressure), 
               fill = "light green") +
  geom_vline(xintercept = mmniBP, size = 1, color = "red") +
  xlim(0, 300)

mheart_rate <- mean(chartevents_tble$heart_rate, na.rm = TRUE)
ggplot(data = chartevents_tble) +
  geom_density(mapping = aes(heart_rate), fill = "light green") +
  geom_vline(xintercept = mheart_rate, size = 1, color = "red")

mrespiratory_rate <- mean(chartevents_tble$respiratory_rate, na.rm = TRUE)
ggplot(data = chartevents_tble) +
  geom_density(mapping = aes(respiratory_rate), fill = "light green") +
  geom_vline(xintercept = mrespiratory_rate, size = 1, color = "red")

mtemp <- mean(chartevents_tble$body_temperature_in_Fahrenheit, na.rm = TRUE)
ggplot(data = chartevents_tble) +
  geom_density(mapping = aes(body_temperature_in_Fahrenheit), 
               fill = "light green") +
  geom_vline(xintercept = mtemp, size = 1, color = "red")
```

## Q7. Putting things together

Let us create a tibble `mimic_icu_cohort` for all ICU stays, where rows are  

- first ICU stay of each unique adult (age at admission > 18)

and columns contain at least following variables  

- all variables in `icustays.csv.gz`  
- all variables in `admission.csv.gz`  
- all variables in `patients.csv.gz`  
- first lab measurements during ICU stay  
- first vital measurements during ICU stay
- an indicator variable `thirty_day_mort` whether the patient died within 30 days of hospital admission (30 day mortality)

```{r}
admissions_tble <- admissions_tble %>% 
                   group_by(subject_id) %>% 
                   arrange(admittime) %>% 
                   distinct(subject_id, .keep_all = TRUE)


mimic_icu_cohort1 <- full_join(icustays_tble, admissions_tble, 
  by = "subject_id")

mimic_icu_cohort2 <- full_join(mimic_icu_cohort1, patients_tble, 
  by = "subject_id")

mimic_icu_cohort3 <- full_join(mimic_icu_cohort2, labevents_tble, 
  by = "subject_id")

mimic_icu_cohort <- full_join(mimic_icu_cohort3, chartevents_tble, 
  by = "subject_id")
```

```{r}
mimic_icu_cohort$thirty_day_mort <- rep(NA, 53150)

for (i in 1:53150) {
  if (is.na(mimic_icu_cohort$deathtime[i])) {
    mimic_icu_cohort$thirty_day_mort[i] <- 0
  }
  else if (mimic_icu_cohort$deathtime[i] - mimic_icu_cohort$admittime[i] > 30) {
    mimic_icu_cohort$thirty_day_mort[i] <- 0
  }
  else {
    mimic_icu_cohort$thirty_day_mort[i] <- 1
  }
}

```

## Q8. Exploratory data analysis (EDA)

Summarize following information using appropriate numerics or graphs.

- `thirty_day_mort` vs demographic variables (ethnicity, language, insurance, marital_status, gender, age at hospital admission)

```{r}
mimic_icu_cohort %>% 
  group_by(ethnicity) %>% 
  summarise(mean = mean(thirty_day_mort))

mimic_icu_cohort %>% 
  group_by(language) %>% 
  summarise(mean = mean(thirty_day_mort))

mimic_icu_cohort %>% 
  group_by(insurance) %>% 
  summarise(mean = mean(thirty_day_mort))

mimic_icu_cohort %>% 
  group_by(marital_status) %>% 
  summarise(mean = mean(thirty_day_mort))

mimic_icu_cohort %>% 
  group_by(gender) %>% 
  summarise(mean = mean(thirty_day_mort))

mimic_icu_cohort %>% 
  group_by(thirty_day_mort) %>% 
  summarise(mean = mean(anchor_age))
  
  
```

- `thirty_day_mort` vs first lab measurements

```{r warning = FALSE}
ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = bicarbonate))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = calcium))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = chloride))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = creatinine))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = glucose))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = hematocrit))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = magnesium))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = potassium))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = sodium))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), 
                               y = white_blood_cell_count))
```

- `thirty_day_mort` vs first vital measurements

```{r warning = FALSE}
ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), 
                               y = systolic_non_invasive_blood_pressure))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), 
                               y = mean_non_invasive_blood_pressure))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = heart_rate))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), y = 
                                 respiratory_rate))

ggplot(data = mimic_icu_cohort) +
    geom_boxplot(mapping = aes(x = as.factor(thirty_day_mort), 
                               y = body_temperature_in_Fahrenheit))
```

- `thirty_day_mort` vs first ICU unit

```{r}
mimic_icu_cohort$first_careunit <- as.factor(mimic_icu_cohort$first_careunit)

mimic_icu_cohort %>% 
  group_by(first_careunit) %>% 
  summarise(mean = mean(thirty_day_mort))
```
